---
title: "Project 1"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Project 1

## Load Libraries

```{r, results=FALSE, message=FALSE, warning=FALSE}
library(RCurl)
library(stringr)
library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
```

## Load Raw Data

Using the `RCurl` library download the raw text data from my GitHub repository. Set the result as `raw_data`
```{r read-data}
file_url <- 'https://raw.githubusercontent.com/nolivercuny/data607/master/project1/tournamentinfo.txt'
raw_data <- getURL(file_url)
```

## Processing

### Converting giant string to dataframe

Get rid of dashes. At this point the data is one very long string.
```{r}
data <- str_replace_all(raw_data,"-", "")
```

Split the data on newlines then on the pipe character
```{r}
split <- data %>% 
  str_split('\n', simplify = TRUE) %>% 
  str_split('\\|', simplify = TRUE)
```

Convert to a dataframe and use the filter function to remove empty rows
```{r}
df <- as.data.frame(split) %>%
  filter(V1 != "", V1 != "\r")
```

### Manipulate the dataframe

At this point we have a dataframe which contains all our data in individual rows but there are two rows representing a single player record. This is because the raw data was structured in such a way that a logical row was actually represented as two newline separated lines in the text file. 

Break up the dataframe into two dataframes so they can be recombined as a single row.
```{r}
df_one <- df %>% filter(row_number() %% 2 == 1) 
df_two <- df %>% filter(row_number() %% 2 == 0)
```
Set the column names to the first row
```{r}
names(df_one) <- df_one[1,]
names(df_one) <- df_one[1,]
```
Merge the two dataframes back into a single data frame. I am also using the `subset` combined with the `select` function to drop two columns that contain no data.
```{r}
data <- bind_cols(df_one,df_two) %>% 
  subset(select=-c(11,22))
```

Drop the first row because it contained the column names and we no longer need that.
```{r}
data <- data[-c(1),]
```

Split column 12 which contains the USCF Id and the Pre and Post ratings for the the player. Then split the pre and post ratings into separate columns as well.
```{r}
data <- separate(data, 12, sep="/", into = c('USCF_ID','PRE_POST'))
data <- separate(data, 13, sep=">", into = c('Pre_Rating','Post_Rating'))

```

Parse the Pre rating column as a number so it can be used in subsequent calculation
```{r}
data$Pre_Rating <- parse_number(data$Pre_Rating)
```

### Computing Opponent's Pre-Rating Mean

1. Loop over every row. 
2. Select the "Rounds" columns (4 through 10). 
3. Because those contain the row number of the opponent parse out that value as a number using `gsub` to replace all non-numeric characters with no value. 
4. Then us `as.numeric` to cast the values as numbers to be used as indexes in the dataframe.
5. Use the opponent index vector to select the opponents pre-ratings.
6. Use the mean function to obtain the average opponent pre-rating with the `na.rm` argument set to `TRUE` in order to remove `NA` values where opponents were not part of the round for the player.
7. Round the mean to the nearest whole value using the `round` function and add that value as a new column in the dataframe.
```{r}
for (i in 1:nrow(data)){
  opponent_indexes <- as.numeric(gsub("[^0-9]","",data[i,4:10]))
  opponents_ratings<-data[opponent_indexes,13]
  rating_average <- mean(opponents_ratings, na.rm = TRUE)
  data$Mean_Opponent_Pre_Rating[i] <- round(rating_average)
}
```

### Clean up for final output
1. Select only relevant columns
2. Remove the `R:` from the player's pre-rating.
3. Rename the columns
4. Trim whitespace
5. Reset the row numbers 
```{r}
final_output <- data[,c(2,3,11,13,23)]
final_output$Pre_Rating <- gsub("R:", "", final_output$Pre_Rating)
names(final_output) <- c('Player’s Name', 'Total Number of Points','Player’s State', 'Player’s Pre-Rating', 'Average Pre Chess Rating of Opponents')
final_output <- final_output %>% 
  mutate(across(where(is.character), str_trim))
rownames(final_output) <- NULL
```


```{r}
kable(final_output,caption="Players",digits = 2, format = "html", row.names = TRUE)
```

```{r}
working_directory <- getwd()
write_csv(final_output,paste(working_directory, "/chess_tournement_players.csv",sep=""))
```