---
title: "Project 2 Wirecutter v Amazon"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Project 2 - The Wirecutter Vs. Amazon

The goal of this analysis is to evaluate the quality of The Wirecutter's product recommendations. I am suspicious that The Wirecutter is bad judge of quality due to my own personal experience of going to purchase a recommended product on Amazon and seeing generally negative sentiment about the product Some of the bad reviews even mentioned The Wirecutter's recommendation as their reason for buying the product.

To perform my analysis I used the article that The Wirecutter posted with recommendations for the best wired earbuds under $200. [https://www.nytimes.com/wirecutter/reviews/the-best-200-in-ear-headphones/](https://www.nytimes.com/wirecutter/reviews/the-best-200-in-ear-headphones/)

## Setup

### Load Libraries
```{r}
library(RCurl)
library(stringr)
library(dplyr)
library(tidyr)
```

### Load Data

```{r}
reviewsUrl <- 'https://raw.githubusercontent.com/nolivercuny/data607/master/project2/datasets/wirecuttervsamazon/Customer%20reviews'
customerReviews <- getURL(reviewsUrl)
fileUrl <- 'https://raw.githubusercontent.com/nolivercuny/data607/master/project2/datasets/wirecuttervsamazon/wirecutter_earbuds.csv'
wirecutterData <- getURL(fileUrl)
```

## Tidy

The Amazon review data was scrapped directly from the product pages for the products reviewed by The Wirecutter. 

```{r split}
customerReviews <- str_replace_all(customerReviews, '\t','')
customerReviewsList <- str_split(customerReviews, '\n')
```


```{r}
spread(customerReviewsList, )
```