---
title: "Homework 5 - Working with XML and JSON in R"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Books 

Here is a table with the 3 books I chose and their properties

|title|authors|page count|publication year|current price on amazon|
|-----|-------|----------|----------------|-----------------------|
|Real World Haskell|Bryan O'Sullivan, John Goerzen, Don Stewart|670|2008|$49.99|
|Programming in Haskell|Graham Hutton|171|2007|$18.97|
|Learn You a Haskell for Great Good!: A Beginner's Guide|Miran Lipovača|400|2011|$42.21|

## JSON

Use the `jsonlite` libraries `fromJSON` function and simply cast as dataframe

```{r results=FALSE, message=FALSE, warning=FALSE}
library(jsonlite)
jsonDf <- as.data.frame(fromJSON("https://raw.githubusercontent.com/nolivercuny/data607/master/homework5/books.json"))
```

## XML

Use the `read_xml` function of the `xml2` library.

Convert the XML to a list. 

Unnest by the root node of the XML

Now you have a structure that looks like the dataframe you want but the values are lists of lists of the actual value due to the XML structure.

Unnest every column two more times to extract the values from the list of lists



```{r results=FALSE, message=FALSE, warning=FALSE}
library(xml2)
library(tidyverse)
xmlDoc <- as_list(read_xml("https://raw.githubusercontent.com/nolivercuny/data607/master/homework5/books.xml"))

# put list into a one column tibble
xmlDf<-as_tibble(xmlDoc) %>%
  unnest_wider(books) %>%
  # unnest same length list cols
  unnest(cols = names(.)) %>%
  # unnest again because each column value is still a list
  unnest(cols = names(.)) %>%
  type_convert()
```


## HTML

Use the web scraping library`rvest`'s `html_table` function and cast the result as a dataframe

```{r results=FALSE, message=FALSE, warning=FALSE}
library(rvest)
htmlDf <- read_html("https://raw.githubusercontent.com/nolivercuny/data607/master/homework5/books.html")

htmlDf <- as.data.frame(html_table(htmlDf))
```

## Are the three data frames identical?

First compare all dataframes and they are not equal because the column names are different.

Set all the column names the same and they are still not identical because the values of each column are different types from each other due to the different ways the libraries are parsing the data into dataframes.

```{r}

(all_equal(xmlDf, htmlDf))
(all_equal(xmlDf, jsonDf))
(all_equal(jsonDf, htmlDf))

colNames <- c("title","authors","page count","publication year","current price on amazon")
names(xmlDf) <- colNames
names(htmlDf) <- colNames
names(jsonDf) <- colNames

(all_equal(xmlDf, htmlDf))
(all_equal(xmlDf, jsonDf))
(all_equal(jsonDf, htmlDf))
```