---
title: "Project 2 West Nile Virus"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Project 2 - West Nile Virus

https://bbhosted.cuny.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_2010109_1&nav=discussion_board_entry&conf_id=_2342994_1&forum_id=_2992508_1&message_id=_54088437_1

https://www.cdc.gov/westnile/statsmaps/cumMapsData.html

There are two tables here with the cases each year in a separate column:

West Nile virus disease cases reported to CDC by state of residence, 1999-2019

West Nile virus neuroinvasive disease cases reported to CDC by state of residence, 1999-2019

The totals per state & per year are already calculated, but the means are yet to be calculated.

The year columns could be pivoted longer and the 2 tables could be combined. Percent of positive cases that developed into neuroinvasive disease could be calculated (per year / per state).

## Setup

### Load Libraries
```{r}
library(RCurl)
library(stringr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(readr)
library(fuzzyjoin)
```

### Load Data

```{r}
westNileCasesUrl <- 'https://raw.githubusercontent.com/nolivercuny/data607/master/project2/datasets/westnile/West-Nile-virus-disease-cases-by-state_1999-2019-P.csv'
westNileCasesRaw <- getURL(westNileCasesUrl)
westNileNeuroUrl <- 'https://raw.githubusercontent.com/nolivercuny/data607/master/project2/datasets/westnile/West-Nile-virus-neuroinvasive-disease-cases-by-state_1999-2019-P.csv'
westNileNeuroRaw <- getURL(westNileNeuroUrl)
```

## Tidy

```{r}

```



