---
title: "Project 2 West Nile Virus"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Project 2 - West Nile Virus



https://www.cdc.gov/westnile/statsmaps/cumMapsData.html

For this analysis I used Victoria McEleney's suggested data sets on the West Nile Virus by state[^1]. 

There are two tables here with the cases each year in a separate column:

West Nile virus disease cases reported to CDC by state of residence, 1999-2019[^2]

West Nile virus neuroinvasive disease cases reported to CDC by state of residence, 1999-2019[^3]

The totals per state & per year are already calculated, but the means are yet to be calculated.

The year columns could be pivoted longer and the 2 tables could be combined. Percent of positive cases that developed into neuroinvasive disease could be calculated (per year / per state).

## Setup

### Load Libraries
```{r}
library(stringr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(readr)
```

### Load Data

```{r}
westNileCasesUrl <- 'https://raw.githubusercontent.com/nolivercuny/data607/master/project2/datasets/westnile/West-Nile-virus-disease-cases-by-state_1999-2019-P.csv'
westNileCasesRaw <- read.csv(westNileCasesUrl, sep = " ", skip = 1)

westNileNeuroUrl <- 'https://raw.githubusercontent.com/nolivercuny/data607/master/project2/datasets/westnile/West-Nile-virus-neuroinvasive-disease-cases-by-state_1999-2019-P.csv'
westNileNeuroRaw <- read.csv(westNileNeuroUrl, sep = " ", skip = 1)
```

## Tidy

```{r}
westNileCasesDf <- westNileCasesRaw %>% 
  as.data.frame()
westNileNeuroDf <- westNileNeuroRaw %>%
  as.data.frame()
```

Fix DC because it's the only one that has three spaces in the name
```{r}
fix_dc <- function(df){
  # District of Columbia 9,78
dc <- df[9,] %>% unite(State,1:3, sep = " ")
df[9,]$State <- dc[1,1]
df[78,]$State <- dc[1,1]

df[9,2:3] <- df[10,1:2]
df[78,2:3] <- df[79,1:2]
df
}
westNileCasesDf<-fix_dc(westNileCasesDf)
westNileNeuroDf <-fix_dc(westNileNeuroDf)
```


```{r}
fix_states <- function(df){
  #there is probably a clever way to find these
  statesWithSpaces <- c(31,33,35,37,39,41,47,49,51,53,61,100,102,104,106,108,110,116,118,120,122,130)
  for (index in statesWithSpaces) {
    dc <- df[index,] %>% unite(State,1:2, sep = " ")
    df[index,]$State <- dc[1,1]
    df[index,2] <- df[index+1,1]
  
  }
  df
}
westNileCasesDf <- fix_states(westNileCasesDf)
westNileNeuroDf <- fix_states(westNileNeuroDf)
```

Filter all the columns with bad/meaningless data. Had to do `gsub` because there were some rows where the number had a comma, which meant it was parsed as `NA` and didn't get filtered
across(where(is.character), str_trim))

```{r}
casesDf <- westNileCasesDf %>% 
  filter(is.na(as.numeric(gsub(",", "",State)))) %>% 
  mutate(across(where(is.character), str_trim))

nueroDf <- westNileNeuroDf %>% 
  filter(is.na(as.numeric(gsub(",", "",State)))) %>% 
  mutate(across(where(is.character), str_trim))

```
Now we have two data sets combined into one. I will need to split the dataframe and recombined it by state.

```{r}
recombine <- function(df) {
  df_one <- df[1:52,]
  df_two <- df[57:109,]
  names(df_two) <- df_two[1,]
  df_two <- df_two[-c(1),]
  bind_cols(df_one,df_two) %>% 
    subset(select = -c(`State...13`))
}
```
```{r}

finalCasesDf <- recombine(casesDf) %>%
  pivot_longer(-State...1) %>%
  mutate(value = as.numeric(gsub(",","",value)))

finalNeuroDf <- recombine(nueroDf) %>% 
  pivot_longer(-State...1) %>%
  mutate(value = as.numeric(gsub(",","",value)))
```

## Analysis

```{r}
sumCases <- finalCasesDf %>%
  group_by(State...1) %>% 
  summarize(m = mean(value)) %>%
  arrange(desc(m))

sumNeuro <- finalNeuroDf %>%
  group_by(State...1) %>% 
  summarize(m = mean(value))%>%
  arrange(desc(m))

final <- inner_join(sumCases, sumNeuro, by = "State...1")

kable(final,caption = "Mean West Nile By State", format = "html") %>% kable_styling("striped") %>% scroll_box(width = "100%")
```

## Conclusions

It appears based on my analysis that the states with the highest population had the highest instances of both West Nile and Neuroinvasive West Nile.

## References

[^1]: https://bbhosted.cuny.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_2010109_1&nav=discussion_board_entry&conf_id=_2342994_1&forum_id=_2992508_1&message_id=_54088437_1

[^2]: West Nile virus disease cases reported to CDC by State of ... (n.d.). Retrieved October 4, 2021, from https://www.cdc.gov/westnile/resources/pdfs/data/West-Nile-virus-disease-cases-by-state_1999-2019-P.pdf. 

[^3]: West Nile virus neuroinvasive disease cases reported to ... (n.d.). Retrieved October 4, 2021, from https://www.cdc.gov/westnile/resources/pdfs/data/West-Nile-virus-neuroinvasive-disease-cases-by-state_1999-2019-P.pdf. 
