---
title: "Homework 4"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---
```{r load-libraries}
library(tidyr)
library(dplyr)
library(RCurl)
library(kableExtra)
```

# Homework 4 - Tidying and Transforming Vaccination Data

The csv contained in my GitHub repository was generated using this website [https://cloudconvert.com/xlsx-to-csv](https://cloudconvert.com/xlsx-to-csv) that converts `.xlsx` files into `.csv` files. 
```{r}
file_url <- 'https://raw.githubusercontent.com/nolivercuny/data607/master/homework4/israeli_vaccination_data_analysis_start.csv'
raw_data <- getURL(file_url)
data <- readr::read_csv(raw_data)
```

## Tidying the data

The Efficacy column was loaded incorrectly and contains a number of distinct values. Separating out using the `separate` function. The column names don't matter at this point so setting them to (A,B,C)
```{r}
data <- data %>% separate(Efficacy, sep = ",", into = c('A','B','C'))
```

Renaming the columns is very manual. The way the data was read in from the CSV it put the column names that spanned multiple subcolumns as the column names into my datafame. Then the sub column names were inserted as the first row.

My goal here was to combine the multi-column names (Population %, Severe Cases) with the descriptive and important sub-column names (Not Vax, Fully Vax). I set the dNames variable to the original names so when I overwrite the column name I can still use the original names.

```{r}
dNames <- names(data)
#Pop % + Not Vax
names(data)[2] <- paste(dNames[2], data[1,2])
#Pop % + Fully Vax
names(data)[3] <- paste(dNames[2], data[1,3])
# Sever Cases + Not Vax per 100k
names(data)[4] <- paste(dNames[3],data[1,4], data[2,4])
# Sever Cases + Fully Vax per 100k
names(data)[5] <- paste(dNames[3],data[1,5], data[2,5])
# vs. severe disease
names(data)[6] <- paste('Efficacy',data[1,6])

```

Remove the first two rows because they are no longer needed

```{r}
data <- data %>% slice(-c(1,2))
```

Now I need to take rows 2 and 4 and add them as the values of two new columns representing the population %

First fill the `Age` column downward
```{r}
data <- data %>% fill(Age, .direction='down')
```
Next split the data into two dataframes so they can be merged back together
```{r}
pop_value_data <- data %>% slice(1,3)
names(pop_value_data)[2] = "Population Not Vax"
names(pop_value_data)[3] = "Population Fully Vax"
pop_percent_data <- data %>% slice(2,4)
```
Last merge them together but only select the first three columns of the second dataframe to prevent duplicate columns
```{r}
final_data <- full_join(pop_value_data, select(pop_percent_data,c(1,2,3)),by="Age")
```


## (1) Do you have enough information to calculate the total population?  What does this total population represent?

**Yes, the populations include both vaccinated and unvaccinated people both over and at or under the age of fifty.**

```{r}
total_population <- sum(as.numeric(gsub(",","",final_data$`Population Not Vax`))) +
  sum(as.numeric(gsub(",","",final_data$`Population Fully Vax`)))
total_population
```
**6,937,546**

## (2) Calculate the Efficacy vs. Disease; Explain your results.

```{r}
#convert the columns to numeric
final_data$`Population Fully Vax`<- as.numeric(gsub(",","",final_data$`Population Fully Vax`))
final_data$`Population Not Vax`<- as.numeric(gsub(",","",final_data$`Population Not Vax`))

final_data$`Severe Cases Not Vax per 100K`<- as.numeric(final_data$`Severe Cases Not Vax per 100K`)
final_data$`Severe Cases Fully Vax per 100K`<- as.numeric(final_data$`Severe Cases Fully Vax per 100K`)


```

```{r}

final_data$`Efficacy vs. severe disease` <- 1 - (
  (final_data$`Population Fully Vax` / 100000 ) * final_data$`Severe Cases Fully Vax per 100K` / final_data$`Population Fully Vax`) /
  (
    (final_data$`Population Not Vax` / 100000 ) * final_data$`Severe Cases Not Vax per 100K` / final_data$`Population Not Vax`) 


```
**My results are wrong but I don't have time to figure out why. I believe based on the data that this shows that the efficacy is high and higher for > 50 year olds than < 50 year olds.**

```{r}
kable(final_data,caption="israeli_vaccination_data_analysis",digits = 2, format = "html", row.names = TRUE)
```


## (3) From your calculation of efficacy vs. disease, are you able to compare the rate of severe cases in unvaccinated individuals to that in vaccinated individuals?


```{r}

```

