---
title: "Week 9 Homework"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Week 9 Homework

In this assignment the goal is to retrieve and load JSON data from the New York Times open data APIs. I chose to use the movie reviews API that they offer to load my data in. The reviews API has a `search.json` route which includes a `critics-pick` query parameter that I used in order to retrieve 20 movie reviews in JSON format.

## Setup

### Load Libraries

For this assignment I only needed two libraries. The `kableExtra` library is used for displaying my data for review purposes. The `jsonlite` library does the heavy lifting of retrieving and parsing the data.

```{r results=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)
library(jsonlite)
```

## Load Raw JSON from NYT

I found the full route for the API by using the New York Times documentation website. My API key is embedded in the URL as a query parameter. 

The `jsonlite` library includes a function called `fromJSON` which not only is able to make the HTTP GET request to the NYT API but also parses the results into a list of data.

```{r}
url <- 'https://api.nytimes.com/svc/movies/v2/reviews/search.json?critics-pick=Y&api-key=FIU5RW8zWaF0jp8Y0w76rUlhusYdmagW'
results <- fromJSON(url)
```

The result of the `fromJSON` request is a 5 item list. 

The first item in the list represents the status of the response I am extracting that value and validating that it equals `"OK"`

The last item in the list is the actual response body already in a dataframe. I simply extract that body as a dataframe and display it in a table.

```{r warning=FALSE}
status <- results[[1]]
print(paste("Response Status: ", status))
if(status != "OK"){
  print("Request failed!")
}else{
 rawDf <- results[[5]]
 rawDf %>% 
   kbl() %>%
     kable_paper(full_width = F) %>%
    scroll_box(width = "100%", height="600px")
}

```

