---
title: "Final Project Data Scraping"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Scraping Articles

## Loading Libraries

I am loading the following libraries

  - tidyverse
    - contains the `rvest` library used for web scraping the Wired.com source data
  - odbc
    - used to connect to Azure SQL server for long-term storage of the data

```{r}
library(tidyverse)
library(rvest)
library(DBI)
library(RCurl)
```

## Web Scraping

We will start with the sitemap. Luckily it is a very simple page which just contains a very long list of links to article archives.
```{r}
siteMapUrl <- "https://www.wired.com/sitemap/"
siteMapHtml <- read_html(siteMapUrl)
```

To select the links to the archives I simply need to select the parent div which wraps the archive list then grab the `a` elements
```{r}
archiveLinkNodes <- siteMapHtml %>%
  html_nodes(".sitemap__section-archive > ul > li > a")
```

I end up with `r length(archiveLinkNodes)` archive links. Each archive contains multiple articles so to simplify the process and decrease the risk of being blocked I will randomly sample the archive list. 

I do want to make that I sample enough data because I want the articles to span a large  time span
```{r}
set.seed(1972)

sampledArchiveLinks <- sample(archiveLinkNodes, 150) %>%
  html_text()
```

Check the years to make sure we have a good sample. It appears the sample includes a link from every year that the Wired.com has published articles online which is an ideal result.
```{r}
sampledArchiveLinks %>%
  stringi::stri_extract_all_regex("(?<=year=)[0-9]+") %>%
  unlist() %>%
  unique() %>%
  parse_number() %>%
  sort()
```

Next I need to use the links to randomly select a subset of articles from the archive. Each link takes the user to another page with a list of articles that were published on the year, month and week listed in the link.

```{r}
links <- list()
for (archiveLink in sampledArchiveLinks) {
  tryCatch(expr={
    siteMapHtml <- read_html(archiveLink)
    archiveLinkNodes <- siteMapHtml %>% 
      html_nodes(".sitemap__section-archive > ul > li > a") %>%
      html_text()
    links[[archiveLink]] <- archiveLinkNodes
  },error=function(e){
    message(paste("Failed to retrieve data for link ", archiveLink))
    message(e)
  },finally = {
    message(paste("Processed URL ", archiveLink))
    Sys.sleep(sample(1:5, 1))
  })
}
write.csv(links, file="final_project_links.csv")
```

```{r}
library(jsonlite)
library(purrr)
library(data.table)

dt_list <- map(links, as.data.table)
dt <- rbindlist(dt_list, fill = TRUE, idcol = T)

dt <- dt %>% rename(source = .id, link = V1)
```


Now we have the links it's time to extract the data

```{r}
extractArticleContent <- function(html) {
  publishDate <- html %>%
  html_node(xpath = '//*[@data-testid="ContentHeaderPublishDate"]') %>% 
  html_text()


author <- html %>%
  html_node(xpath = '//*[@data-testid="BylineName"]') %>% 
  html_text()


subject <- html %>%
  html_node(xpath = '//*[@data-testid="ContentHeaderHed"]') %>% 
  html_text()


byLine <- html %>%
  html_node(xpath = '//*[@data-testid="ContentHeaderAccreditation"]/div') %>% 
  html_text()

category <- html %>% 
  html_nodes(".rubric__link") %>%
  html_text()

articleBody <- html %>% 
  html_node(".body__inner-container") %>% 
  html_text()

c(publishDate, author, subject, byLine, category, articleBody)
}
```

```{r}
articlesDf <- data.frame(publishDate = character(), author = character(), subject = character(), byLine = character(), category = character(), body = character(), sourceUrl = character(), retrievedAt = character())
```

```{r}
for (link in links[1:25]) {
    for(article in sample(link,10)){
      tryCatch(expr={
        html <- read_html(article)
        data <- extractArticleContent(html)
        articlesDf <- articlesDf %>% 
        add_row(publishDate = data[1], author = data[2], subject = data[3], byLine = data[4], category = data[5], body = data[6], sourceUrl = article, retrievedAt =format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"))
      },error=function(e){
        message(paste("Failed on article ", article))
        message(e)
      },finally = {
        message(paste("Processed URL:", article))
        Sys.sleep(sample(1:5, 1))
      })
    }
}
```

Now that I have the data from a large amount of articles spanning a large time range, I will store that data in a database so it can be cleaned and analyzed

backup as csv just in case
```{r}
write.csv(articlesDf, file = "articles.csv")
```

## Database

Establish connection
```{r}

#For testing use in memory DB
# con <- dbConnect(RSQLite::SQLite(), dbname = ":memory:")

#Uncomment to manipulate real DB
con <- dbConnect(RSQLite::SQLite(), dbname = "final_project.db")
```

```{r}
dbWriteTable(con,"LinksXref",dt, append=TRUE)
```

Insert records
```{r}
dbWriteTable(con,"Articles",articlesDf, append=TRUE)
```

Make sure our data was inserted correctly
```{r}
dbGetQuery(con, "SELECT * FROM Articles")
```

```{r}
dbDisconnect(con)
```

